Sample Length at Which Block Bootstrapping is Effective for Estimation of Parameter of Time Series

### 2023-01-26 Manuscript/manuscript.tex
Another update

@@

For all simulation studies, but $\mu$ and $\sigma_{x}$ can be kept constant, but increasing temporal dependence factors such as AR and MA Coefficients can make block bootstrap less effective in estimation, so its impact on the method's performance should be analyzed. For a certain variation of an AR(1) process, block bootstrap 95\% confidence intervals for each of 3 parameters ($\mu$, $\sigma_{x}$, and the autocorrelation function$\rho$) can be replicated for 1000 simulated samples of a certain length. For each individual confidence interval, it can be recorded whether or not the interval includes the true parameter. The coverage rate, or the proportion of intervals that include the true parameter, can then be recorded. If the parameter is being properly estimated by the block bootstrap, the coverage rate should reflect this by being close to the confidence level of 95\%. If this is not the case, the method is not working as well as it should. The width of the confidence interval for each replication is dependent on the variability of the parameter of the time series. If the interval is on average not wide enough to capture the parameter at the expected rate of 95\%, it indicates that the variability of the parameter is underestimated. If the interval is too wide and captures the parameter at a rate above 95\%, it indicates that the variability of the parameter is overestimated. Since the coverage rate is a proportion based on 1000 independent binary outcomes (the interval recovers the parameter or it doesn't), a 95\% confidence interval of the coverage rate for each parameter can also be created. The inclusion of .95 in the interval indicates that the block bootstrap method is working well for its respective parameter. If this is not the case for this sample length, the procedure can be conducted again with a larger sample length. This can be repeated until the method is determined to be working acceptably, meaning that .95 is included in the 95\% confidence interval of the coverage rate of the parameter.

1000 replicates of length-n random samples of an AR(1) process with theoretical $\mu$ = 0, $\sigma_{x}$ = 1, and $\theta$ = 0.1 were simulated using R. For each of three target parameters ($\mu$ = 0, $\sigma_{x}$ = 1, and $\theta$ = 0.1, three types of block bootstrap (with block length l) confidence intervals (percentile, empirical, and BCA) were constructed for each replicate sample. The statistic used to estimate $\mu$ is the sample mean $\bar{x}$, the statistic used to estimate $\sigma$ is the sample standard deviation $S^2$, and the statistic used to estimate $\theta$ is the auto-correlation function of the sample. For each of these nine categories of confidence intervals, the coverage rate of its respective parameter was computed, along with a 95\% confidence interval of the proportion given 4000 trials. For a certain variation of the process, the study was repeated for n = 125, 1000, 3375, and 8000. The only parameter of the process that was varied was $\theta$, and 0.1, 0.3, and 0.5 were the values used.

@@

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.2}
  \caption{At $\phi = .2$, all intervals covered $\mu$ correctly at $n = 300$ and $n = 400$. The alternate percentile interval and the $\hat{\theta}_{n}$-centered percentile interval cover $\phi$ well at $n = 500$. The $\bar\theta_n^*$-centered BCA interval slightly under-covers $\phi$ for all sample lengths observed.}
  \label{fig:plot_.2}
\end{figure}

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.4}
  \caption{}
  \label{fig:plot_.4}
\end{figure}

>>>>

The only parameter of the AR(1) process that was varied was $\phi$, as the temporal dependence affects the performance of block bootstrap. $\phi$ = 0 (a standard normal distribution), 0.2, and 0.4 were used in this study.

@@

If $\theta$ is not covered at that $n$, a larger sample size must be necessary. $n = 100, 200, 300, 400, and 500$ were used in this study.

%For all simulation studies, but $\mu$ and $\sigma_{x}$ can be kept constant, but increasing temporal dependence factors such as AR and MA Coefficients can make block bootstrap less effective in estimation, so its impact on the method's performance should be analyzed. For a certain variation of an AR(1) process, block bootstrap 95\% confidence intervals for each of 3 parameters ($\mu$, $\sigma_{x}$, and the autocorrelation function$\rho$) can be replicated for 1000 simulated samples of a certain length. For each individual confidence interval, it can be recorded whether or not the interval includes the true parameter. The coverage rate, or the proportion of intervals that include the true parameter, can then be recorded. If the parameter is being properly estimated by the block bootstrap, the coverage rate should reflect this by being close to the confidence level of 95\%. If this is not the case, the method is not working as well as it should. The width of the confidence interval for each replication is dependent on the variability of the parameter of the time series. If the interval is on average not wide enough to capture the parameter at the expected rate of 95\%, it indicates that the variability of the parameter is underestimated. If the interval is too wide and captures the parameter at a rate above 95\%, it indicates that the variability of the parameter is overestimated. Since the coverage rate is a proportion based on 1000 independent binary outcomes (the interval recovers the parameter or it doesn't), a 95\% confidence interval of the coverage rate for each parameter can also be created. The inclusion of .95 in the interval indicates that the block bootstrap method is working well for its respective parameter. If this is not the case for this sample length, the procedure can be conducted again with a larger sample length. This can be repeated until the method is determined to be working acceptably, meaning that .95 is included in the 95\% confidence interval of the coverage rate of the parameter.

%1000 replicates of length-n random samples of an AR(1) process with theoretical $\mu$ = 0, $\sigma_{x}$ = 1, and $\theta$ = 0.1 were simulated using R. For each of three target parameters ($\mu$ = 0, $\sigma_{x}$ = 1, and $\theta$ = 0.1, three types of block bootstrap (with block length l) confidence intervals (percentile, empirical, and BCA) were constructed for each replicate sample. The statistic used to estimate $\mu$ is the sample mean $\bar{x}$, the statistic used to estimate $\sigma$ is the sample standard deviation $S^2$, and the statistic used to estimate $\theta$ is the auto-correlation function of the sample. For each of these nine categories of confidence intervals, the coverage rate of its respective parameter was computed, along with a 95\% confidence interval of the proportion given 4000 trials. For a certain variation of the process, the study was repeated for n = 125, 1000, 3375, and 8000. The only parameter of the process that was varied was $\theta$, and 0.1, 0.3, and 0.5 were the values used.

@@

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.2}
  \caption{At $\phi = .2$, all intervals covered $\mu$ correctly at $n = 300$ and $n = 400$. The alternate percentile interval and the $\hat{\theta}_{n}$-centered percentile interval cover $\phi$ well at $n = 500$. The $\bar\theta_n^*$-centered BCA interval slightly under-covers $\phi$ for all sample lengths observed. The alternate BCA interval under-covers $\phi$, and its performance deteriorates as $n$ increases, as does the $\bar\theta_n^*$-centered percentile interval but with lower coverage rates and more noticeable deterioration. The $\hat{\theta}_{n}$-centered BCA interval significantly under-covers $\phi$, although there is no clear deterioration.}
  \label{fig:plot_.2}
\end{figure}

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.4}
  \caption{At $\phi = .4$,}
  \label{fig:plot_.4}
\end{figure}
